### 2022-08-15 main/chapter_2ljets.tex
tweaks for Chris' excellent feedback; fix y label on drhigh n0 plot

This chapter follows an unusual order.
A more ordinary ordering would first present preparations made in the search
project --- blind to data ---
followed later by the results --- unblinded.
Instead, this chapter first reports the unblinded results and follows those
with in-depth discussions of individual elements, partially in light of those
results.
This unconventional ordering makes sense for two reasons.
Firstly, it avoids iterating over those elements twice, once before unblinding
and again after;
search project has many little and loosely-related pieces that I attempt to
to cover locally to reduce the reader's need to jump between far-flung corners
of the chapter.
Secondly, and sadly, many design decisions can only be explained in light of
the unblinded results, since major aspects of the search were reworked in the
nineteen months between its unblinding and eventual publication.

@@

To test non-standard (supersymmetric) models against observation
we add signal contributions on top of the modelled backgrounds.
We use two signal models in the $\twoljets$-electroweak analysis.
One is an MSSM chargino-neutralino model (C1N2) that acts to produce
$\chargino_1\textrm{--}\neutralino_2$ pairs (via off-shell $W^\pm$ resonances),
that decay via $W\rightarrow q\bar q$ and $Z\rightarrow \ell^\pm \ell^\mp$
respectively to produce two leptons, jets, and two $\neutralino_1$ LSPs.
The other is a GMSB model which, from a variety of initial states, produces
$\neutralino_1\textrm{--}\neutralino_1$ pairs that decay via Higgs and $Z$
bosons, in variable branching fractions,
to final states again containing $\ell^\pm \ell^\mp$, hadronic partons,
and invisible gravitinos.
Diagrams illustrating these C1N2 and GMSB signal models are shown in
Figure~\ref{fig:2ljets_signal_diagrams}.
Example effects of these signal models are displayed in signal regions of the
$\twoljets$-electroweak search are displayed in
Figure~\ref{fig:2ljets_signal_examples}.
All of these regions require two leptons, jets, and
missing transverse momentum, and the signal expectations would be added on top
of the backgrounds if Nature realized these supersymmetric models.

@@

Both signal models have two free parameters that our search varies.
At each parameter point of a signal model, it is added to the background model
and tested against data in the $\cls$ prescription.
Contours showing which parameters of those signal models are labelled as
excluded are displayed for
C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
GMSB in~\ref{fig:2ljets_contours_gmsb}.
In both of these figures, space to the lower left of the contours is
labelled as excluded, and space on the other side of the observed limit contour
is not excluded.
The observed contours are somewhat larger than expected due to observed
deficits in data.

% exclusion plots
\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption[
Contours for the C1N2 model in the $\twoljets$-electroweak search
]{%
Contours for the C1N2 model in the $\twoljets$-electroweak
search~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded, and its dotted
neighbours show alternative results that are obtained if all signal
cross-sections are varied up and down by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the distribution of the test statistics.
Grey areas are observed limits from the two-lepton results from previous
\atlas\ searches~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\cls$ prescription
with asymptotic approximations~\cite{Cowan:2010js}.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
\caption[
Contours for the GMSB model in the $\twoljets$-electroweak search
]{%
Contours for the GMSB model in the $\twoljets$-electroweak
search~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded, and its dotted
neighbours show alternative results that are obtained if all signal
cross-sections are varied up and down by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the distribution of the test statistics.
Grey areas are observed limits from the two-lepton results from previous
\atlas\ searches~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\cls$ prescription
with asymptotic approximations~\cite{Cowan:2010js}.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_gmsb}
\end{figure}

Signal regions of the $\twoljets$-electroweak analysis are combined to form
five discovery regions.
Results of their model-independent interpretations are presented in
Table~\ref{tab:2ljets_discovery}.
The same data deficits drive stronger bounds on possible signals that manifest
as smaller signal expectations limits $S_{\mathrm{obs}}^{95}$.

% upper limits
% definitely want this after pictures
\FloatBarrier
\begin{table}[tp]
\centering
% custom separator for aligned +-
% https://stackoverflow.com/a/2132998
\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
& \multicolumn{2}{c}{Fitted}
& Data
& $A\epsilon\sigma_{\mathrm{obs}}^{95}\,\mathrm{fb}$
& $S_{\mathrm{obs}}^{95}$
& $S_{\mathrm{exp}}^{95}$
& $\clb$
& $p_{s\!=\!0}$
\\[1.5ex]
DR-OffShell & $22.1$ & $2.7$ & $21$ & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$
\\[.5ex]
DR-Low & $22$ & $4$ & $18$ & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$
\\[.5ex]
DR-Int & $35$ & $4$ & $38$ & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$
\\[.5ex]
DR-High & $3.9$ & $0.5$ & $0$ & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0$ & $0.50$
\\[.5ex]
DR-$\llbb$ & $0.51$ & $0.20$ & $0$ & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$
\\[.5ex]
\end{tabular*}
\caption[
Upper limit results in $\twoljets$-electroweak discovery regions
]{%
Upper limit results in $\twoljets$-electroweak discovery
regions~\cite{atlas2022searches}.
Each fitted yield is in the background-only model constrained by the data in
all control regions and the discovery region.
Limits are intended to reflect constrains on additive signal contributions.
\\[0.5em]
Left to right:
the region name,
the post-fit background expectation,
the number of data observed,
the observed $95\%$ $\cls$ upper limit on the visible cross-section
$\langle\epsilon\sigma\rangle_\mathrm{obs}^{95}$,
its corresponding signal expectation $S_\mathrm{obs}^{95}$,
the expected $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
as would be obtained were the test statistic given by its central or
$\pm1$-sigma variations,
$\clb$ evaluated with the signal expectation set to the observed upper limit,
and the discovery $p$-value (capped at $0.5$) with its equivalent significance.
\\[0.5em]
Upper limits use the one-sided profile likelihood test
statistic~\cite{Cowan:2010js}.
The discovery $p$-value uses a profile likelihood test statistic in a one-sided
test.
All $p$-values are estimated by ``toy'' simulations of alternative data.
Values with errors are rounded to one or two significant figures according
the PDG rounding scheme~\cite{pdg2020review}.%
}
\label{tab:2ljets_discovery}
\end{table}

@@

\clearpage
\FloatBarrier
\section{Summaries}
\label{sec:2ljets_context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact ---
this search project exists to supplement those results with information from
the full LHC Run~2 data-set that was collected between in the years $2015$
through to $2018$ and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$.
Leveraging the increased data quantity is one key contribution of $\twoljets$.
Analysis quality also might be improved if our designs take lessons from
the experiences of previous work.

@@

\subsubsection{Doubts}
Disappearing excesses have encouraged some doubts about RJR variables in
certain social circles.
The existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06},
which approximates the $3$ lepton result of~\cite{atlas_rjr_23l_SUSY_2017_03}
with non-RJR ``laboratory-frame'' event variables,
is public evidence of this.

@@

A lack of reliable simulations motivates reliance on `data-driven' background
estimates in RJR selections, which, from my perspective, are hard to interpret
or trust.
Indeed,  all `data-driven' methods rely not only on data but on
hypothetical extrapolations that may not be well motivated, particularly when
studying complicated non-linear event variables such as RJR variables.
Although RJR variables can appear to offer appealing separations of
supersymmetric signals from Standard Model backgrounds,
these practical issues motivate my decision to use any RJR event variables in
the $\twoljets$-electroweak analysis.

@@

The $\twoljets$-strong search targets strongly interacting supersymmetric
effects in cases where their decays include pairs of charged leptons.
Like the RJR and electroweak $\twoljets$ searches, signal models targeted by
the $\twoljets$-strong search produce jets and invisible LSPs.
Unlike the RJR and electroweak parts, however, not all of these leptons are
produced by $Z$ boson decays;
as illustrated in Figure~\ref{fig:2ljets_strong_signal_diagrams}, decay trees
featuring sleptons can also produce lepton pairs.
Lepton pairs produced in these slepton interactions are of the same flavour
and opposite charge, like those from $Z$ boson decays, but they have different
mass dilepton mass characteristics.

@@

To target its signal models, the $\twoljets$-strong search defines seven
overlapping signal regions.
Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to $m(Z)$,
and the others are binned in $\mll$ for sensitivity to the various
kinematic edges across the grid of signal model parameters.

The largest Standard Model background to this search is dileptonic $\ttbar$,
partly because no $b$-tagging veto is applied.
All three signal models produce same-flavour lepton pairs, but $\ttbar$
produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
equal rates ---
the $\twoljets$ search exploits this difference in flavour symmetry to derive
a data-driven background estimate for all of its flavour-symmetric backgrounds,
and have its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
The background estimate works by taking different-flavour data, weighting them
by factors to calibrate for differences in the leptons' reconstruction, and
histogramming them in same-flavour regions.
Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.

@@

Personally, I find these jagged contours hard to interpret.
If I had a pet theory that landed on one of these sharp
region-transition edges, what should I make of these data?
I wouldn't know.
It seems certain that some combination of the regions could have combined their
information to produce a continuous and more informative contour.
Where model predictions change continuously in these signal spaces, it seems
that our inferences should also.
To aid interpretations and to allow justifiable combinations of regions'
information, I therefore chose to design the $\twoljets$-electroweak with only
orthogonal regions, and without any switching of regions between models.

@@

Worse, I disregarded the popular idea that every region should have a large
background expectation, preferably of ten events or more.
This requirement is typically justified by the limitations of approximate
statistical methods that work
``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2014}.
But the Poisson likelihood works for infinitesimally small
expectations~\cite{skilling2009cosmology}, and in the small-bin limit one has
continuous distributions with Poisson normalizations
(sometimes called ``extended'' in high-energy
physics~\cite{barlow1990extended})
that are standard elsewhere in high energy physics, and are easily analysed
by the standard likelihood fitting methods.

Events appearing in a region with minimal background is evidence in favour
of new contributions, and that only strengthens as the background approaches
zero.
For example, the top quark was observed by \dzero\ using seven regions all with
background expectations of $1$ or less, for a total of $3.79\pm0.55$ expected
and $17$ observed~\cite{abachi1995observation}.
If they could do that with technology from the year $1995$, we can do it now.
This is a fact, but only if we can really trust our predictions of zero
backgrounds.

@@

Unfortunately, the $\twoljets$-electroweak regions designs were finalized
before unblinding and before I appreciated this problem with background
uncertainties in the smaller regions.
Various dubious workaround are invoked to assign plausible variations despite
small yields, which we shall in the remainder of this chapter.
As seen in the splash summary of Section~\ref{sec:2ljets_splash}, the
$\twoljets$-electroweak search observes agreement with post-fit backgrounds
except for some local deficits of data below backgrounds.
Exclusion contours exceed their expected range because of these deficits,
but they can only be trusted if the background model is trusted.

@@

For the C1N2 model, these public Run~2 \atlas\ searches explore the
all-hadronic~\cite{SUSY-2018-41},
$2/3\ell$ compressed~\cite{atlas_susy_compressed_2l_2018_run2},
$3\ell$ on-shell~\cite{SUSY-2018-06}, and
$3\ell$/same-sign~\cite{SUSY-2018-09} channels.
In comparison to these other searches, our exclusion contour of
Figure~\ref{fig:2ljets_contours_c1n2} excludes a unique area of parameter
space within $m(\chargino_1,\neutralino_2) \in 550\textrm{--}700\,\eV[G]$
and $m(\neutralino_1) \in 200\textrm{--}300\,\eV[G]$.
The all-hadronic search is most sensitive at the highest parent masses,
and three-lepton data of in the on-shell and compressed searches are most
sensitive at intermediate and small (off-shell/compressed) mass splittings.

@@

From its name, it is clear that $\twoljets$ analyses features of events with
two charged leptons (those are electrons or muons --- not taus),
and at least one jet.
Though not in the name, a $\met$ requirement is also implicit from the context
of $R$-parity conserving supersymmetric models.
This section serves to describe how events are selected for the $\twoljets$
analyses, and which event variables are extracted for use in its selections.

@@

\subsection{Trigger}
\label{sec:2ljets_trigger}
There would be no event variables if not for the trigger reducing the data to
a manageable size.
When living our lives as humans, we are bombarded with sensory data and must
pay attention only to those few gems of any use, to avoid exhausting our
processing capacity.
Running \atlas\ is no different.

@@

In total, we use $14$ different two-lepton trigger selections, each of which
selects for an electron pair, a muon pair, or an emu (electron-muon pair).
To describe those triggers, we use $\pt^{\ell_i}$ for the transverse momentum
of the $i\mathrm{th}$ hardest lepton.
All triggers require $\pt^{\ell_i}$ values to exceed thresholds in
the range $8\textrm{--}26\,\eV[G]$.
To detail all trigger requirements and their variations with data period
would be too tedious and not enlightening, but their $\pt^{\ell_i}$ requirements
are summarized here.
Please do not read them all in detail.
\begin{itemize}
\item Electron pair: requirements vary between
\begin{itemize}
\item $\pt^{e_2} > 12\,\eV[G]$ and
\item $\pt^{e_2} > 17\,\eV[G]$.
\end{itemize}
\item Muon pair: various requirements include
\begin{itemize}
\item $\pt^{\mu_2} > 10\,\eV[G]$,
\item $\pt^{\mu_2} > 14\,\eV[G]$,
\item $\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
\item $\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
\end{itemize}
\item Emu pair: various asymmetric requirements include
\begin{itemize}
\item $\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
\item $\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
\item $\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
\end{itemize}
\end{itemize}
Throughout the data used in $\twoljets$, trigger requirements
have increased in hardness with time.
The point to listing these requirements is to show trigger activations are
non-trivial.
This motivates lepton selections which exceed these trigger thresholds,
to avoid the need to model their jagged activations and variable
efficiencies near to those thresholds~\cite{
atlas_trigger_egamma_run2,
atlas_trigger_muon_run2
}.
To avoid these thresholds, all leptons selected for the $\twoljets$-electroweak
regions are required to have $\pt > 25\,\eV[G]$.

@@

The benefit of an additional trigger could be in increased signal yields if
that trigger does not also harmfully increase inseparable backgrounds.
A study, \emph{performed by Knut Vadla}, investigated the effects
of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $10\%$;
improvement. Adding one-lepton triggers was found to increase efficiency by
only a few percentage points, depending on the model.
The $\twoljets$ team considered that to be too little to be worthwhile.

>>>>

This chapter follows an unusual order.
A more ordinary ordering would first present preparations made in the search
project --- blind to data ---
followed later by the results --- unblinded.
Instead, this chapter first reports the unblinded results and follows those
with in-depth discussions of individual elements, partially in light of those
results.
This unconventional ordering makes sense for two reasons.
Firstly, it avoids iterating over those elements twice, once before unblinding
and again after;
this search project has many little and loosely-related pieces that I attempt
to cover locally to reduce the reader's need to jump between far-flung corners
of the chapter.
Secondly, and sadly, many design decisions can only be explained in light of
the unblinded results, since major aspects of the search were reworked in the
nineteen months between its unblinding and eventual publication.

@@

To test non-standard (supersymmetric) models against observation
we add signal contributions on top of the modelled backgrounds.

The $\twoljets$-electroweak search uses two signal models,
which we name C1N1 and GMSB and illustrate with diagrams in
Figure~\ref{fig:2ljets_signal_diagrams}.
One is an MSSM chargino-neutralino model (C1N2) that acts to produce
$\chargino_1\textrm{--}\neutralino_2$ pairs (via off-shell $W^\pm$ resonances),
that decay via $W\rightarrow q\bar q$ and $Z\rightarrow \ell^\pm \ell^\mp$
respectively to produce two leptons, jets, and two $\neutralino_1$ LSPs.
The other is a GMSB model which, from a variety of initial states, produces
$\neutralino_1\textrm{--}\neutralino_1$ pairs that decay via Higgs and $Z$
bosons, in variable branching fractions,
to final states again containing $\ell^\pm \ell^\mp$, hadronic partons,
and invisible gravitinos.
Example effects of these signal models are displayed in signal regions of the
$\twoljets$-electroweak search are displayed in
Figure~\ref{fig:2ljets_signal_examples}.
All of these regions require two leptons, jets, and
missing transverse momentum, and the signal expectations would be added on top
of the backgrounds if Nature realized these supersymmetric models.

@@

Both signal models have two free parameters that our search varies.
At each parameter point of a signal model, the search proceed by evaluating
event rates (yields) from that signal model with those parameters,
and considering what would happen of those event yields were added to a
Standard Model background.
Each signal-plus-background model is then tested against observed data in the
$\cls$ prescription.
From these tests, contours showing which parameters of those signal models are
labelled as excluded are derived, and displayed in
Figure~\ref{fig:2ljets_contours_c1n2} for C1N2 and
in~\ref{fig:2ljets_contours_gmsb} for GMSB.
In both of these figures, space to the lower left of the contours is
labelled as excluded, and space on the other side of the observed limit contour
is not excluded.
The observed contours are somewhat larger than expected due to observed
deficits in data that we have shown in Figure~\ref{fig:2ljets_summary}.

% exclusion plots
\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption[
Contours for the C1N2 model in the $\twoljets$-electroweak search
]{%
Contours for the C1N2 model in the $\twoljets$-electroweak
search~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded, and its dotted
neighbours show alternative results that are obtained if all signal
cross-sections are varied up and down by factors intended to describe the size
theoretical uncertainties.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the distribution of the test statistics.
Grey areas are observed limits from the two-lepton results from previous
\atlas\ searches~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\cls$ prescription
with asymptotic approximations~\cite{Cowan:2010js}.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
\caption[
Contours for the GMSB model in the $\twoljets$-electroweak search
]{%
Contours for the GMSB model in the $\twoljets$-electroweak
search~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded, and its dotted
neighbours show alternative results that are obtained if all signal
cross-sections are varied up and down by factors intended to describe the size
theoretical uncertainties.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the distribution of the test statistics.
Grey areas are observed limits from the two-lepton results from previous
\atlas\ searches~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\cls$ prescription
with asymptotic approximations~\cite{Cowan:2010js}.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_gmsb}
\end{figure}

Signal regions of the $\twoljets$-electroweak analysis are combined to form
five discovery regions.
Results of their model-independent interpretations are presented in
Table~\ref{tab:2ljets_discovery}.
The same data deficits drive stronger bounds on possible signals that manifest
as smaller limits on the signal expectations, which are labelled
$S_{\mathrm{obs}}^{95}$ in the figure.

% upper limits
\begin{table}[tp]
\centering
% custom separator for aligned +-
% https://stackoverflow.com/a/2132998
\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
& \multicolumn{2}{c}{Fitted}
& Data
& $A\epsilon\sigma_{\mathrm{obs}}^{95}\,\mathrm{fb}$
& $S_{\mathrm{obs}}^{95}$
& $S_{\mathrm{exp}}^{95}$
& $\clb$
& $p_{s\!=\!0}$
\\[1.5ex]
DR-OffShell & $22.1$ & $2.7$ & $21$ & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$
\\[.5ex]
DR-Low & $22$ & $4$ & $18$ & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$
\\[.5ex]
DR-Int & $35$ & $4$ & $38$ & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$
\\[.5ex]
DR-High & $3.9$ & $0.5$ & $0$ & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0$ & $0.50$
\\[.5ex]
DR-$\llbb$ & $0.51$ & $0.20$ & $0$ & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$
\\[.5ex]
\end{tabular*}
\caption[
Upper limit results in $\twoljets$-electroweak discovery regions
]{%
Upper limit results in $\twoljets$-electroweak discovery
regions~\cite{atlas2022searches}.
Each fitted yield is in the background-only model constrained by the data in
all control regions and the discovery region.
Limits are intended to reflect constrains on additive signal contributions.
\\[0.5em]
Left to right:
the region name,
the post-fit background expectation,
the number of data observed,
the observed $95\%$ $\cls$ upper limit on the visible cross-section
$A\epsilon\sigma_{\mathrm{obs}}^{95}\,\mathrm{fb}$,
its corresponding signal expectation $S_\mathrm{obs}^{95}$,
the expected $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
as would be obtained were the test statistic given by its central or
$\pm1$-sigma variations,
$\clb$ evaluated with the signal expectation set to the observed upper limit,
and the discovery $p$-value capped at $0.5$.
The cross-section and signal yield limits are related by a factor of our
integrated luminosity $139.0\,\mathrm{fb}^{-1}$.
\\[0.5em]
Upper limits use the one-sided profile likelihood test
statistic~\cite{Cowan:2010js}.
The discovery $p$-value uses a profile likelihood test statistic in a one-sided
test.
All $p$-values are estimated by ``toy'' simulations of alternative data.
Values with errors are rounded according the PDG rounding
scheme~\cite{pdg2020review}.%
}
\label{tab:2ljets_discovery}
\end{table}

@@

\clearpage
\FloatBarrier
\section{Summaries}
\label{sec:2ljets_context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact ---
this search project exists to supplement those results with information from
the full LHC Run~2 data-set that was collected between in the years $2015$
through to $2018$ and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$.
Leveraging the increased data quantity is one key contribution of $\twoljets$.
How well those data are used depends on how well the search designed, so
$\twoljets$ further benefits from lessons learned from the experiences of
past analyses.

@@

\subsubsection{Doubts}
Disappearing excesses have encouraged some doubts about RJR variables in
certain social circles;
the existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06},
which approximates the $3$-lepton result of~\cite{atlas_rjr_23l_SUSY_2017_03}
with non-RJR ``laboratory-frame'' event variables,
is public evidence of this.

@@

A lack of reliable simulations motivates reliance on `data-driven' background
estimates in RJR selections, which, from my perspective, are hard to interpret
or trust.
Indeed,  all `data-driven' methods rely not only on data but on
hypothetical extrapolations that may not be well motivated, particularly when
studying complicated non-linear event variables such as RJR variables.
Although RJR variables can appear to offer appealing separations of
supersymmetric signals from Standard Model backgrounds, these practical issues
motivate my decision to not use any RJR event variables in the
$\twoljets$-electroweak analysis.

@@

The $\twoljets$-strong search is designed to target strongly interacting
supersymmetric effects in cases where their decays include pairs of charged
leptons.
Like the RJR and electroweak $\twoljets$ searches, signal models targeted by
the $\twoljets$-strong search produce jets and invisible LSPs.
Unlike the RJR and electroweak parts, however, not all of these leptons are
produced by $Z$ boson decays;
as illustrated in Figure~\ref{fig:2ljets_strong_signal_diagrams}, decay trees
featuring sleptons can also produce lepton pairs.
Lepton pairs produced in these slepton interactions are of the same flavour
and opposite charge, like those from $Z$ boson decays, but they have different
mass dilepton mass characteristics.

@@

To target its signal models, the $\twoljets$-strong search defines seven
overlapping signal regions.
Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to $m(Z)$,
and the others are binned in $\mll$ for sensitivity to the various
kinematic endpoints that arise across the grid of various signal model
parameters.

The largest Standard Model background to this search is dileptonic $\ttbar$,
partly because no $b$-tagging veto is applied.
All three signal models produce same-flavour lepton pairs, but $\ttbar$
produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
equal rates.
The $\twoljets$ search exploits this difference in flavour symmetry to derive
a data-driven background estimate for all of its flavour-symmetric backgrounds,
and have its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
The background estimate works by taking different-flavour data, weighting them
by factors to calibrate for differences in the leptons' reconstruction, and
histogramming them in same-flavour regions.
Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.

@@

Personally, I find these jagged contours hard to interpret.
If I were interested in a theory that landed on one of these sharp
region-transition edges, what should I make of these data?
I wouldn't know.
It seems certain that some combination of the regions could have combined their
information to produce a continuous and more informative contour.
Where model predictions change continuously in these signal spaces, it seems
that our inferences should also.
To aid interpretations and to allow justifiable combinations of regions'
information, I therefore chose to design the $\twoljets$-electroweak with only
orthogonal regions, and without any switching of regions between models.

@@

Worse, I disregarded the popular idea that every region should have a large
background expectation, preferably of ten events or more.
This requirement is typically justified by the limitations of approximate
statistical methods that work
``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2014}.
But the Poisson likelihood works for infinitesimally small
expectations~\cite{skilling2009cosmology}, and in the small-bin limit one has
continuous of `unbinned' distributions with Poisson normalizations
(sometimes called ``extended'' in high-energy
physics~\cite{barlow1990extended})
that are standard elsewhere in high-energy physics,
and are easily analysed by the standard likelihood fitting
methods~\cite{cowan1998statistical}.
Examples uses of unbinned modelling include analyses of hadronic
resonances in
\lhcb~\cite{Tullythesis,lhcb2018563}, searches for exotic effects in
\atlas~\cite{EXOT-2018-08},
and notably the observation of the
Higgs boson~\cite{HIGG-2012-27}.

Events appearing in a region with minimal background is evidence in favour
of new contributions, and that only strengthens as the background approaches
zero.
For example, the top quark was observed by \dzero\ using seven regions all with
background expectations of one event or less, for a total of $3.79\pm0.55$
expected and seventeen observed~\cite{abachi1995observation}.
If they could do that with technology from the year $1995$, we can do it now.
This is a fact, but only if we can really trust our predictions of zero
backgrounds.

@@

Unfortunately, the $\twoljets$-electroweak regions designs were finalized
before unblinding and before I appreciated this problem with background
uncertainties in the smaller regions.
Various workaround are invoked to assign plausible variations despite
small yields and failing statistical precision, which shall be seen in the
remainder of this chapter.
As seen in the splash summary of Section~\ref{sec:2ljets_splash}, the
$\twoljets$-electroweak search observes agreement with post-fit backgrounds
except for some local deficits of data below backgrounds.
Exclusion contours exceed their expected range because of these deficits,
but they can only be trusted if the background model is trusted.

@@

For the C1N2 model, these public Run~2 \atlas\ searches explore the
all-hadronic~\cite{SUSY-2018-41},
$2\ell/3\ell$ compressed~\cite{atlas_susy_compressed_2l_2018_run2},
$3\ell$ on-shell~\cite{SUSY-2018-06}, and
$3\ell$/same-sign~\cite{SUSY-2018-09} channels.
In comparison to these other searches, our exclusion contour of
Figure~\ref{fig:2ljets_contours_c1n2} excludes a unique area of parameter
space within $m(\chargino_1,\neutralino_2) \in 550\textrm{--}700\,\eV[G]$
and $m(\neutralino_1) \in 200\textrm{--}300\,\eV[G]$.
The all-hadronic search is most sensitive at the highest parent masses,
and three-lepton data of in the on-shell and compressed searches are most
sensitive at intermediate and small (off-shell/compressed) mass splittings.

@@

The `$\twoljets$' name is intended to indicate interest in events with two
charged leptons (those are electrons or muons --- not taus), and at least one
jet.
Though not in the name, an $\met$ requirement is also implicit from the context
of $R$-parity conserving supersymmetric models.
This Section~\ref{sec:2ljets_events} serves to describe how events are selected
for the $\twoljets$ analyses, and which event variables are extracted for use
in its selections.

@@

\subsection{Trigger}
\label{sec:2ljets_trigger}
There would be no event variables without events, and \atlas\ only records
when triggered.
When living our lives as humans, we are bombarded with sensory data and must
pay attention only to those few gems of any use, to avoid exhausting our
processing capacity.
Running \atlas\ is no different --- its triggering is necessary to reduce the
bombardment of collision data to a manageable size.

@@

In total, we use $14$ different two-lepton trigger selections, each of which
selects for an electron pair, a muon pair, or an emu (electron-muon pair).
To describe those triggers, we use $\pt^{\ell_i}$ for the transverse momentum
of the $i\mathrm{th}$ hardest lepton.
All triggers require $\pt^{\ell_i}$ values to exceed thresholds in
the range $8\textrm{--}26\,\eV[G]$.
To detail all trigger requirements and their variations with data period
would be too tedious and not enlightening, but their $\pt^{\ell_i}$ requirements
are summarized here.
Please do not read them all in detail ---
the point of this listing is to show trigger activations are varied and
non-trivial.
\begin{itemize}
\item Electron pair: requirements vary between $\pt^{e_2} > 12\,\eV[G]$ and
$\pt^{e_2} > 17\,\eV[G]$.
\item Muon pair: various requirements include
$\pt^{\mu_2} > 10\,\eV[G]$,
$\pt^{\mu_2} > 14\,\eV[G]$,
$\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
$\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
\item Emu pair: various asymmetric requirements include
$\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
$\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
$\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
\end{itemize}
Trigger requirements have increased in hardness with time throughout the data
used in $\twoljets$.

This complicated collection of trigger requirements motivates lepton selections
that exceed these trigger thresholds,
such that they avoid the need to accurately model their jagged activation
thresholds and the variable efficiencies near to those thresholds~\cite{
atlas_trigger_egamma_run2,
atlas_trigger_muon_run2
}.
To avoid the most of these thresholds, all leptons selected for the
$\twoljets$-electroweak regions are required to have $\pt > 25\,\eV[G]$.

@@

The benefit of an additional trigger could be in increased signal yields if
that trigger does not also harmfully increase inseparable backgrounds.
A study, \emph{performed by Knut Vadla}, investigated the effects
of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $10\%$;
improvement. Adding one-lepton triggers was found to increase efficiency by
only a few percentage points, depending on the model.
We in the $\twoljets$ team considered that to be too little to be worthwhile.
